Modal Abstractions for Virtualizing Memory Addresses

### 2023-04-04 programlogic.tex
+

 One possible approach is to abstract away the physical page-table-walk pointsto ($\textsf{L}_{4}\_\textsf{L}_{1}\_\textsf{PointsTo}$ in Figure \ref{fig:strongvirtualpointsto}) from the virtual pointsto relation via replacing it with read-only abstract map entries -- a ghost (abstract) map entry per \textsf{L}$_4$\_\textsf{L}$_{1}$ mappings -- which can be used as a resource ($ \sumwalkabs\vaddr\qfrac\paddr$) whose ownership ensures that the virtual address ($\vaddr$) is mapped to a page (referenced by $\paddr$) ,and as a token to access to the physical page-table-walk pointsto (\textsf{L}$_4$\_\textsf{L}$_{1}$) stored in an address-space invariant ($\mathcal{I}$\textsf{ASpace} in Figure \ref{fig:peraspaceinvariant}). We use \textsf{Iris}'s authorative ghost-map logical machinery
\[\mathcal{A}\textsf{bsPTableWalk} \stackrel{\triangle}{=} \textsc{Auth} (\; \mathcal{W}_{64} \;\rightarrow_{\textrm{fin}} \;  (\mathcal{W}_{64} \; \rightarrow_{\textrm{fin}}\;  (\textsc{Frac }, \mathord{+}) \times (\textsc{Agree } \Loc,\mathord{=}) ))\]
to realize these reasoning principles ($\mathcal{A}\textsf{bsPTableWalk}$).
\begin{remark}[Exchanging Page-Table-Walk-Summarization Mappings (Tokens) with Physical Page-Table-Walk Pointstos ]
  $\mathcal{A}\textsf{PTableWalk}$'s fragmental-ownership represents an individual key-value pair (per-virtual-address) which can be handled as a \textit{look-up} resource for an individual virtual address.
  To build more intuition, we can think of the use-case \textit{unmapping-pages} as collecting the fragmental map pairs, e.g. $\sumwalkabs\vaddr\qfrac\paddr$, from collecting the assertions, removing them from the ghost map, and then updating the page tables.
  Critically, physical full ownership ($\ownGhost\gammaPred{\authfull{\ptableabswalk\ptablestore}}$) of the page tables, as a \textit{singleton resource}, lives in the central invariant keeping physical-page-table pointstos per-virtual address translation.
  With this mechanism we can separate the virtual-to-physical translation from the physical memory involved, and consequently
  \begin{itemize}
  \item virtual pointsto assertions, whose definition is shown in Figure \ref{fig:virtualpointstosharing}, can refer to their physical memory assertions representing their page-address translation in the physical-page-tables ($\textsf{L}_{4}\_\textsf{L}_{1}\_\textsf{PointsTo}$) which reside in the address-space invariant ($\mathcal{I}$\textsf{ASpace} in Figure \ref{fig:peraspaceinvariant}) using individual key-value pair, i.e. instances of ($\sumwalkabs\vaddr\qfrac\paddr$)
  \item unlike the case in the strong virtual-pointsto relation in Figure \ref{fig:strongvirtualpointsto}, physical-page-table-walk pointsto assertions ($\textsf{L}_{4}\_\textsf{L}_{1}\_\textsf{PointsTo}$) for all virtual addresses are moved into the address-space invariant ($\mathcal{I}$\textsf{ASpace} in Figure \ref{fig:peraspaceinvariant}), and resides with authorative ownership of $\mathcal{A}\textsf{PTableWalk}$ ($\ownGhost\gammaPred{\authfull{\ptableabswalk\ptablestore}}$) which respects all entries of the abstract virtual-address translation mappings ($\sumwalkabs\vaddr\qfrac\paddr$).
    \end{itemize}
\end{remark}
\begin{figure*}
\[
\begin{array}{l}
    \vaddr\mapsto_{\textsf{v}}\{\textsf{q}\}\vpage \stackrel{\triangle}{=} 
    \exists \entryo \ldotp
    \sumapaces\crthree\delta \ast 
  \sumwalkabs\vaddr\qfrac\paddr \ast 
   \nfpointsto{\mask\vaddr\ft\entryo}{\mask\vaddr\tw\entryo}\vpage\qfracone\naddr
\end{array}
\]
\caption{Virtual-Pointsto for Sharing Pages}
  \label{fig:virtualpointstosharing}
\end{figure*}
%\subsection{Address-Space Abstraction}
%\label{sec:aspacemodal}

>>>>

  One possible approach is to abstract away the physical page-table-walk pointsto ($\textsf{L}_{4}\_\textsf{L}_{1}\_\textsf{PointsTo}$ in Figure \ref{fig:strongvirtualpointsto}) from the virtual pointsto relation. To do so, we introduce the protocol separating read-only (fragmental) and authorative (full ownership) accesses to the page tables, specifically $\textsf{L}_{4}\_\textsf{L}_{1}\_\textsf{PointsTo}$. We realize this protocol by embedding it into an already existing \textsf{Iris}'s ghost-map (or view\_maps) construction
  \[\mathcal{A}\textsf{bsPTableWalk} \stackrel{\triangle}{=} \textsc{Auth} (\; \mathcal{W}_{64} \;\rightarrow_{\textrm{fin}} \;  (\mathcal{W}_{64} \; \rightarrow_{\textrm{fin}}\;  (\textsc{Frac }, \mathord{+}) \times (\textsc{Agree } \Loc,\mathord{=}) ))\]
 where
  \begin{itemize}
  \item we decorate the fragmental ownership of the ghost-map ($ \sumwalkabs\vaddr\qfrac\paddr$) with the physical $\textsf{L}_{4}\_\textsf{L}_{1}\_\textsf{PointsTo}$ whose ownership just ensures that the virtual address ($\vaddr$) is mapped to a page (referenced by $\paddr$)
  \item we place the authorative ownership $\mathcal{A}\textsf{PTableWalk}$ of the ghost-map in an address-space invariant ($\mathcal{I}$\textsf{ASpace} in Figure \ref{fig:peraspaceinvariant}) whose ownership grants sound update on the page table entries
  \end{itemize}

\begin{remark}[Exchanging Page-Table-Walk-Summarization Mappings (Tokens) ]
  $\mathcal{A}\textsf{PTableWalk}$'s fragmental-ownership represents an individual key-value pair (per-virtual-address) which can be handled as a \textit{look-up} resource for an individual virtual address.
  To build more intuition on the convenience of this protocol, we can think of the use-case \textit{unmapping-pages} as collecting the fragmental map pairs, e.g. $\sumwalkabs\vaddr\qfrac\paddr$, from collecting the assertions, removing them from the ghost map, and then updating the page tables.
\end{remark}
\begin{figure*}
\[
\begin{array}{l}
    \vaddr\mapsto_{\textsf{v}}\{\textsf{q}\}\vpage \stackrel{\triangle}{=} 
    \exists \entryo \ldotp
    \sumapaces\crthree\delta \ast 
  \sumwalkabs\vaddr\qfrac\paddr \ast 
   \nfpointsto{\mask\vaddr\ft\entryo}{\mask\vaddr\tw\entryo}\vpage\qfracone\naddr
\end{array}
\]
\caption{Virtual-Pointsto for Sharing Pages}
  \label{fig:virtualpointstosharing}
\end{figure*}
%\subsection{Address-Space Abstraction}
%\label{sec:aspacemodal}
